---
title: "Week3 assignment - question 2"
date: "Feb 9th, 2025"
author: "Taegeon Yu"
---

## Replicating the Python codes from question 1 into R codes

```{r}
## read and manipulate the dataset
df <- readxl::read_xls("clinics.xls")
df <- df[, c("bizState", "locLat", "locLong")]
colnames(df) <- c("state", "latitude", "longitude")
head(df)
```

* The same sub-data used in Python code is created.

```{r}
# ensure the latitude and longitude are numeric values
df$latitude <- as.numeric(df$latitude)
df$longitude <- as.numeric(df$longitude)
```

```{r}
# haversine function
haversine <- function(lat1, lon1, lat2, lon2) {
  MILES <- 3959  
  # convert degrees into radians for code compatability
  lat1 <- lat1 * pi / 180
  lon1 <- lon1 * pi / 180
  lat2 <- lat2 * pi / 180
  lon2 <- lon2 * pi / 180
  
  dlat <- lat2 - lat1
  dlon <- lon2 - lon1
  
  a <- sin(dlat / 2)^2 + cos(lat1) * cos(lat2) * sin(dlon / 2)^2
  c <- 2 * asin(sqrt(a))
  
  return(MILES * c)
}
```
* The haversine function equivalent to that of the Python code is created.
* We will use for-loop, apply, and vectorization techniques to compare with Python's computation efficiency.

```{r}
for_loop_function <- function(df) {
  distances <- numeric(nrow(df))
  for (i in seq_len(nrow(df))) {
    distances[i] <- haversine(40.671, -73.985, df$latitude[i], df$longitude[i])
  }
  return(distances)
}
```

* The for loop function similar to that of the Python code is created.

```{r}
apply_function <- function(df) {
  return(apply(df, 1, function(row) {
    haversine(40.671, -73.985, as.numeric(row["latitude"]), as.numeric(row["longitude"]))
  }))
}
```

* The apply function similar to that of the Python code is created.

```{r}
vect_function <- function(df) {
  return(haversine(40.671, -73.985, df$latitude, df$longitude))
}
```

* The vectorization function similar to that of the Python code is created.

```{r}
# compare three methods in terms of efficiency
library(microbenchmark)

benchmark <- microbenchmark(
  for_loop = for_loop_function(df),
  apply_method = apply_function(df),
  vectorized = vect_function(df),
  times = 100
)
```

```{r}
# print the computation time as "ms"
print(benchmark, unit = "ms")
```
* The results are out, but we only need the computation time with mean values.

```{r}
results <- summary(benchmark)

df_results <- data.frame(
  Method = c("For Loop", "Apply", "Vectorized"),
  # convert to ms
  Execution_Time_ms = results$mean / 1000000  
)

df_results
```
* The execution time for for-loop, apply, and vectorization function is tabulated.
* It looks like all the functions used for comparison shows that R clearly outperforms Python for efficiency for this case.
* The apply function, which took the longest time in R, shows significantly better efficiency than Python's numpy vectorization.
