---
title: "Week 1 Assignment"
date: "Jan 27th, 2025"
author: "Taegeon Yu"
---

## Question 1

```{r}
# importing the data.table library to fast read datasets
library(data.table)

# importing the dplyr library to use pipeline
library(dplyr)

# importing the tidyr library to drop null values
library(tidyr)
```

```{r}
# reading the patient data
patient <- fread("Downloads/healthcare-data/Patient.csv")
```

```{r}
# reading the mortality data
mortality <- fread("Downloads/healthcare-data/Mortality.csv")
```

```{r}
# converting the MISSING value into null value
patient <- patient %>% 
  mutate(Gender = na_if(Gender, "MISSING"))

# dropping null values
patient <- patient %>% 
                    drop_na()

# subseting the Gender variable into male and female
patient <- patient %>% 
              filter(Gender %in% c("male", "female"))

patient
```

```{r}
# merging the datasets (patient, mortality) on Patient ID
patient_mort <- merge(patient, mortality, by = "PatientID", all.x = TRUE)
patient_mort
```

```{r}
# creating a new variable with patient's death status
patient_mort <- patient_mort %>% 
                mutate(PatientStatus = ifelse(is.na(DateOfDeath), "Alive", "Dead"))
```

```{r}
# creating a subset data with gender and patient status
gender_death <- patient_mort[,.(Gender, PatientStatus)]
gender_death
```

```{r}
# calculating the death rate of each gender

gender_death[, .(
  MortalityRate = sum(PatientStatus == "Dead") / .N
), by = Gender]
```

-   The death rate of male patients (approx 36.2%) is slightly higher than that of female patients (approx 35.1%).

## Question 2

```{r}
# reading the outpatient dataset
outpatient <- fread("Downloads/healthcare-data/OutpatientVisit.csv")
outpatient
```

```{r}
# reading the disease map dataset
disease_map <- fread("Downloads/healthcare-data/DiseaseMap.csv")
disease_map
```

```{r}
# sorting the ICDs
setorder(outpatient, ICD10_1, ICD10_2, ICD10_3)
setkey(disease_map, ICD10)
```

```{r}
# melting the ICDs for better compatability
outpatient_melt <- melt(outpatient, 
                        id.vars = c("PatientID"),
                        measure.vars = c("ICD10_1", "ICD10_2", "ICD10_3"),
                        variable.name = "ICD_Type",
                        value.name = "ICD10")

outpatient_melt
```

-   The ICD10_1,2,3 are now combined into ICD_Type with corresponding ICD10 values.

```{r}
# subseting the disease map data
dmap <- disease_map[, .(ICD10, Condition)]
dmap
```

```{r}
# merging outpatient_melt and dmap datasets
pattern_df <- merge(outpatient_melt, dmap, by = "ICD10")
# dropping the duplicated rows
pattern_df <- pattern_df[!duplicated(pattern_df)]
pattern_df
```

```{r}
# merging the pattern_df and gender_df datasets
gender_df <- patient[, .(PatientID, Gender)]
pattern_df <- merge(pattern_df, gender_df, by = "PatientID")
pattern_df
```

-   The dataset for question 2 is ready after few processing and merging.

```{r}
# calculating the total number of unique male and female patients
total_patients_gen <- pattern_df[, .(
  MaleTotal = uniqueN(PatientID[Gender == "male"]),
  FemaleTotal = uniqueN(PatientID[Gender == "female"]),
  All = uniqueN(PatientID)
)]

total_patients_gen
```

-   The total number of male and female are 5897 and 6310 respectively with total number of 12207.
-   Now let's look at the number of patients by gender in every condition.

```{r}
# calcularting the counts of unique patients by gender and condition
patient_condition <- pattern_df[, .(
  Male = uniqueN(PatientID[Gender == "male"]),
  Female = uniqueN(PatientID[Gender == "female"]),
  All = uniqueN(PatientID)
), by = Condition]

patient_condition
```

-   The number of patients in terms of condition, the hyper tension has the highest number of patients.
-   Now let's convert these numbers into percentage.

```{r}
# converting the numbers to percentage
patient_condition <- patient_condition[, `:=`(
  Male = round((Male / total_patients_gen$MaleTotal) * 100, 2),
  Female = round((Female / total_patients_gen$FemaleTotal) * 100, 2),
  All = round((All / total_patients_gen$All) * 100, 2)
)]

# adding the percentage sign
patient_condition <- patient_condition[, `:=`(
  Male = sprintf("%.1f%%", Male),
  Female = sprintf("%.1f%%", Female),
  All = sprintf("%.1f%%", All)
)]
```

-   The number of each patients for condition are divided by total number of gender and percentage signs were added to the values.

```{r}
# final table
patient_condition
```

-   Patients were most invulnerable to hypertension, showing approximately 40 percent in general, followed by obesity with approximately 22 percent.

## Question 3

```{r}
# extracting the year from visit date and death date into new variables
outpatient[, VisitYear:= year(as.Date(VisitDate))]
mortality[, DeathYear := year(as.Date(DateOfDeath))]
```

```{r}
# finding the visiting year for every patient
first_visits <- outpatient[, .(FirstVisitYear = min(VisitYear)), by = PatientID]
first_visits
```

```{r}
# merging the first_visits and mortality datasets on patient id
date_data <- merge(first_visits, mortality, by = "PatientID", all.x = TRUE)

# dropping the third column from the merged dataset and dropping the null values
date_data <- date_data[, -3]
date_data <- date_data %>% 
                drop_na()
```

```{r}
date_data
```

```{r}
# setting the data table with year range from 2005 to 2018
year <- data.table(Year = 2005:2018)
year
```

```{r}
# calculating the number of patients at risk (people who had their first visit on or before the specific year and whose year of death is after or in the specific year.)
year[, RiskPopulation := sapply(Year, function(year) {
  date_data[FirstVisitYear <= year & DeathYear >= year, .N]
})]
```

```{r}
year
```

```{r}
# calculating the number of dead patients
year[, Deaths := sapply(Year, function(year){
  date_data[DeathYear == year, .N]
})]
```

```{r}
# calculating the mortality rate according to patients at risk and deaths
year[, MortalityRate := round(ifelse(RiskPopulation > 0, Deaths/RiskPopulation, 0), 3)]
year
```

-   The patients at risk and deaths keep increasing until 2011 and reduces after that year.
-   The mortality rate keeps increasing, reaching to 100 percent in 2018.

```{r}
library(ggplot2)

# plotting the mortality rate
ggplot(data = year, aes(x = Year, y = MortalityRate)) + 
  geom_line(color = "blue", size = 1) +
  labs(title = "Mortality Rate Patterns (2005 ~ 2018)",
      x = "Mortality Rate", y = "Year") + 
  scale_x_continuous(breaks = seq(2005, 2018, by = 1)) +
  theme_minimal()
```

-   As indicated in the final table, the mortality rate keeps increasing.

```{r}
ggplot(data = year, aes(x = Year, y = RiskPopulation)) + 
  geom_line(color = "red", size = 1) +
  labs(title = "Risk Population Patterns (2005 ~ 2018)",
      x = "Risk Population", y = "Year") + 
  scale_x_continuous(breaks = seq(2005, 2018, by = 1)) +
  theme_minimal()
```

-   As it was indicated in the final table as well, the risk population keep increasing until year 2011 and drops starting from 2012.
